<!-- 
---
title: "RStudio exercise 3: Logistic Regression"
author: "Hans Hellén"
date: "Wed Nov 22 23:14:59 2017"
output: html_document
---
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE))
```

#RStudio exercise 3: Logistic Regression

## Data Wrangling
For details and links to further information about the data have a look at the wrangling script [create_alc.R](https://github.com/daphnia/IODS-project/blob/master/data/create_alc.R) at my [repository](https://github.com/daphnia/IODS-project). The wrangled data is [here](https://github.com/daphnia/IODS-project/blob/master/data/student_alc_analysis.csv).

##Analysis

The data is about student alcohol consumption researched among students in mathematics and Portugal language courses. The variables are explained here at the [source of data](https://archive.ics.uci.edu/ml/datasets/Student+Performance). Let's first read the data set into memory and have a look at the variable names in the data set. 
```{r echo=T}
#getwd()
#NOTE: some functions are commented out to make the knitted output more simple.
alc <- read.csv("data/student_alc_analysis.csv", header = TRUE)
library(dplyr)
glimpse(alc)
```